---
title: "IE360: Homework 2"
author: "Barış Arslan 2019402129"
date: "11/5/2022"
output: html_document
---

<style type="text/css">
  title {
  font-style: bold}
  body{
  font-family: Times New Roman;
  font-size: 12pt;
}
</style>

## Required Packages
```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(urca)
library(data.table)
library(forecast)
library(ggplot2)
library(GGally)
library(corrplot)
library(zoo)
library(lubridate)
```

## 1. Introduction

In the Homework 2 of IE360: Statistical Forecasting and Time Series course, the ultimate task is to forecast the sales of UGS for every quarter of 2007 by using time series
regression methods.

Here are the first and last six rows of the data:
```{r echo=FALSE, message=FALSE, warning=FALSE}
data <- read.csv(file = "IE360_Spring22_HW2_data.csv")
data<-data%>%
  rename(UGS=`Unleaded.Gasoline.Sale..UGS.`,
         RNUV=`RNUV`,
         NLPG=`X..LPG.Vehicles..NLPG.`,
         PU=`Price.of.Unleaded.Gasoline..PU.`,
         PG=`Price.of.Diesel.Gasoline..PG.`,
         NUGV=`X..Unleaded.Gasoline.Vehicles..NUGV.`,
         NDGV=`X..of.Diesel.Gasoline.Vehicles..NDGV.`,
         GNPA=`GNP.Agriculture`,
         GNPC=`GNP.Commerce`,
         GNP=`GNP.Total`)

data$Quarter <- as.yearqtr(gsub(pattern = "_Q", replacement = "q", data$Quarter),format="%Yq%q")
data$UGS <- as.numeric(gsub(pattern = " ", replacement = "", data$UGS))
data$NLPG <- as.numeric(gsub(pattern = " ", replacement = "", data$NLPG))
data$NUGV <- as.numeric(gsub(pattern = " ", replacement = "", data$NUGV))
data$GNPA <- as.numeric(gsub(pattern = " ", replacement = "", data$GNPA))
data$GNPC <- as.numeric(gsub(pattern = " ", replacement = "", data$GNPC))
data$GNP <- as.numeric(gsub(pattern = " ", replacement = "", data$GNP))

#trend
data<-data%>%mutate(trend=row_number())
#seasonality
data$season2 <- ifelse(quarters(data$Quarter) == "Q2", 1, 0)
data$season3 <- ifelse(quarters(data$Quarter) == "Q3", 1, 0)
data$season4 <- ifelse(quarters(data$Quarter) == "Q4", 1, 0)
head(data)
tail(data)
```
   

## 2. Time Series Plot of UGS
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=6, fig.width=12}
ggplot(data[1:28,1:2], aes(Quarter, UGS, group=1)) +
  geom_line(data = data[1:28,1:2], size=1, color="red", linetype = "solid") + geom_point(color="red", linetype = "solid") +
   ggtitle("Unleaded Gasoline Sales (UGS) over 2000-2006 (Quarterly)") +
    xlab("Quarters") + ylab("Unleaded Gasoline Sales (UGS)") +
    theme_linedraw()
```

In the time series plot of UGS, there is a decreasing trend and an obvious seasonality for every year. In addition, the time series does not have stationary mean and variance. 
KPSS test is used to check the stationarity of the series. Since the test statistic is higher than critical values, the time series is **not stationary**.


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
unt_test=ur.kpss(data$UGS) 
summary(unt_test)

tsUGS = ts(data$UGS,freq=7)
ugs_decom = decompose(tsUGS)
plot(ugs_decom, col=3)
```



## 3. Autocorrelation Function of UGS
```{r echo=FALSE, message=FALSE, warning=FALSE}
acf(data$UGS[1:28], plot = FALSE)
acf(data$UGS[1:28], plot = TRUE, main="Autocorrelation Function of the Unleaded Gasoline Sales (UGS)")
```

The time series is significantly autocorrelated on **lag 1 and 4**. 


## 4. Model Selection

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
ggpairs(data.frame(data[,2:11]), cardinality_threshold = 32) + ggtitle("Correlation Diagram")
```

## 4.1. Alternative Model I

In correlation plots of UGS, there can be seen a significant correlation between UGS and other parameters NLPG, NUGV, GNPA.
That is the reason why the regression modeling will start with those as predictors.

```{r echo=FALSE, message=FALSE, warning=FALSE}
fit1<-lm(formula = UGS~NLPG+NUGV+GNPA, data=data[1:28,])
summary(fit1)
plot(fit1)
```


## 4.2. Alternative Model II

In this alternative model, trend and seasonality factors are added to the model. 
Those factors are defined and added to the data as "trend", "season2", "season3" and "season4" columns.

```{r echo=FALSE, message=FALSE, warning=FALSE}
data.frame(data[,1],data[,12:15])
fit2<-lm(formula = UGS~NLPG+NUGV+GNPA+trend+season2+season3+season4, data=data[1:28,])
summary(fit2)
plot(fit2)
```

## 4.3. Alternative Model III
Since NUGV and GNPA are not significant to the model, they are removed from the model.

```{r echo=FALSE, message=FALSE, warning=FALSE}

fit3<- lm(formula = UGS~NLPG+trend+season2+season3+season4, data=data[1:28,])
summary(fit3)
plot(fit3)
```

## 4.4. Alternative Model IV
UGS’s logarithm is taken to linearize it and used as target variable in the models. Likewise, variables with nonlinear increase can be included in the models as a predictor by taking the logarithm. In addition, data number 6, 12 and 28 are outliers and they are removed from the model.
This is the best model because "Residual standard error" and "p-value" are close to 0, "Adjusted R-squared" is close to 1.
**Alternative Model IV will be used in forecasting of UGS in 2007.**

```{r echo=FALSE, message=FALSE, warning=FALSE}
fit4<-lm(formula = log(UGS)~log(NLPG)+trend+season2+season3+season4, data=data[(1:28)!=6 & (1:28)!=12 & (1:28)!=28,])
summary(fit4)
plot(fit4)
```


## 5. Model Validation
```{r echo=FALSE, message=FALSE, warning=FALSE}
checkresiduals(fit4)
```
In residual analysis it can be seen that residuals are not autocorrelated and distributed normally. Also residuals scatter around mean zero and seem to have a constant variance. Thus this model satisfies the assumptions and it is good enough to use for the prediction of the next year's quarterly sales.



## 6. Forecasts for Unleaded Gasoline Sales for 2007
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=6, fig.width=12}
copy <- data
prediction <- exp(predict(fit4, copy))
data$Prediction_UGS <- prediction

forecast = data.frame(data$Quarter, data$UGS, data$Prediction_UGS)
forecast

ggplot(forecast) +
  geom_line(aes(x=data.Quarter, y=data.UGS,color="Actual"),size=1)+
  geom_line(aes(x=data.Quarter, y=data.Prediction_UGS,color="Forecasted"),size=1)+
  labs(x="Quarters",
       y="Sales",
       title="Actual and Forecasted Values of UGS")+
  theme_linedraw()

```

## 7. Conclusion

In conclusion, final forecasting model results are as follows.
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=6, fig.width=12}
print(paste("The predicted value of 2007-Quarter 1 : ", data$Prediction_UGS[29]))
print(paste("The predicted value of 2007-Quarter 2 : ", data$Prediction_UGS[30]))
print(paste("The predicted value of 2007-Quarter 3 : ", data$Prediction_UGS[31]))
print(paste("The predicted value of 2007-Quarter 4 : ", data$Prediction_UGS[32]))
```

## Additional Notes
My RMarkdown file is located at the progress journal. [Click here](https://bu-ie-360.github.io/spring22-BarisArslan/HW2/HW2_Baris_Arslan.Rmd)


